---
title: "Reproducible Research Week 2 Project"
author: "Cody Coleman"
date: "8/21/2019"
#output: html_document
output: md_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(Hmisc)
library(data.table)
```
________________________________________________________
## Loading and Preprocessing the Data

#### 1. Load the data

```{r, LoadData}
if(!file.exists('activity.csv')){
  if(!file.exists('activity.zip')){
    download.file('https://github.com/CodyColeman/RepData_PeerAssessment1', destfile = 'activity.zip', mode = 'wb')
  }
  unzip('activity.zip')
}
activity_data <- read.csv('activity.csv')
```

#### 2. Quick data exploration

```{r, DataSneakPeek}
names(activity_data)
head(activity_data)
summary(activity_data)
str(activity_data)
```

#### 3. Create a few variables

```{r, Variables}
steps <- activity_data$steps
date <- activity_data$date
interval <- activity_data$interval
```
________________________________________________________
## What is the mean total number of steps taken per day?

#### 1. Make a histogram of the number of steps per day. 
```{r, Hist&Mean}
steps_per_day <- tapply(steps, date, sum, na.rm = TRUE)
hist(steps_per_day, xlab = 'Total Steps (per day)', ylab = 'Frequency', breaks = 20, col = 'steelblue')
```

#### 2. Give the mean and median of the total number of steps per day.
```{r, MeanMedian}
mean_steps <- mean(steps_per_day)
median_steps <- median(steps_per_day)

print(paste0('The mean number of steps per day is ',  mean_steps))
print(paste0('The median number of steps per day is ',  median_steps))
```
________________________________________________________
## What is the average daily activity pattern?

#### 1. Creat a time series plot for average steps per 5 minute increment.
```{r, TimeSeriesPlot}
avgsteps_per_interval <- aggregate(x = steps, by = list(interval), FUN = mean, na.rm = TRUE)
colnames(avgsteps_per_interval) <- c('Intervals', 'Steps')
ggplot(avgsteps_per_interval, aes(x = Intervals, y = Steps)) + geom_line() + xlab('Intervals (5 minute increments)') + ylab('Total Steps per Interval')
```

#### 2. Give the 5-minute interval that, on average, contains the maximum number of steps. 

```{r, MaxStepsOnAvg}
maxSteps <- avgsteps_per_interval[which(avgsteps_per_interval$Steps == max(avgsteps_per_interval$Steps)), ]
maxSteps
print(paste0('Thus, on average, in the ', maxSteps$Intervals, 'th interval, there are ', round(maxSteps$Steps, digits = 3), ' steps taken.' ))
```
________________________________________________________
## Imputing missing values

#### 1. Describe a strategy for imputing missing data. 

1. We will view the total number of missing data to be sure that there is, in fact, missing data. 
2. We will replace all missing data with the avereage number of steps taken without the missing data. 

#### 2. Show the strategy for imputing missing data. 

###### 1. Total number of missing data
```{r, NAcount}
#This number can also be seen in the summary of the original data.
NAcount <- sum(is.na(activity_data))
NAcount
```

###### 2. Create new dataset.

```{r, Imputation}
activity_data_impute <- activity_data
activity_data_impute$steps <- impute(activity_data$steps, fun = mean)
```

#### 3. Create Histogram of total number of seps taken each day after missing values are imputed. 

```{r, HistImputed}
steps_per_day_imp <- tapply(activity_data_impute$steps, activity_data_impute$date, FUN = sum)
hist(steps_per_day_imp, xlab = "Steps per day (Imputed)", ylab = 'Frequency', col = 'violetred', breaks = 20)
```

#### 4. Give new mean and median for imputed data. 

```{r, newMean/Median}
newMeanSteps <- mean(steps_per_day_imp)
newMedianSteps <- median(steps_per_day_imp)

print(paste0('The new mean number of steps after imputation per day is ',  newMeanSteps))
print(paste0('The new median number of steps after imputation per day is ',  newMedianSteps))
```
________________________________________________________

## Are there differences in activity patterns between weekdays and weekends?

#### 1. Seperate Weekends from Weekdays.

```{r}
activity_data_impute$date <- as.Date(activity_data_impute$date)
activity_data_impute$weekday <- weekdays(activity_data_impute$date)
activity_data_impute$weekend <- ifelse(activity_data_impute$weekday == 'Saturday' | activity_data_impute$weekday == 'Sunday', 'Weekend', 'Weekday')
```

#### 2. Create a panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends.

```{r, PanelPlotWeekends/days}
aggregatedActivityImp <- aggregate(steps ~ interval + weekend, data = activity_data_impute, mean)

ggplot(aggregatedActivityImp, aes(interval, steps, color = weekend)) + geom_line() +
facet_grid(weekend ~ . ) +
xlab('Interval (5 minutes)') +
ylab('Average Steps')  
ggtitle("Weekday vs. Weekend Activities")
```
